---
title: "SOC 6100 Fall 2018 Assignment 03 R Notebook"
author: "Malcolm S. Townes"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
    github_document: default
    html_notebook: default
    always_allow_html: yes
---

## Introduction
This is an R Notebook for Assignment 03 in SOC 6100 Fall 2018.

## Project Set Up
The following code chunk enables the R Notebook to integrate seemlessly with the project organization format. This is normally included in the R Notebook to simplify file calls and enable file portability but it has been causing an error. Per Chris Prener of Saint Louis University, the error is generated because the `here::here()` function has not been tested with certain combinations of functions. To work around this problem, I've embedded the `here()` function where I enter a file path when necessary.

```{r setup-project}
knitr::opts_knit$set(root.dir = here::here())
```

## Load Dependencies
The following code chunk loads package dependencies required to perform the necessary tasks. Basic tasks include importing, reading, wrangling, and cleaning data; selecting a subset of the data; checking for unique observations, and analyzing missing data. 
```{r load-dependencies}
library(tidyverse) # loads the basic R packages
library(here) # enables file portability
library(readr) # functions for reading data
library(dplyr) # functions for data wrangling
library(janitor) # functions for data cleaning
library(naniar) # functions for analyzing missing data
library(ggplot2) # functions for data visualizations
```

## Load Raw Data
The following code chunk loads the raw data for Assignment02, which was the cleaned data from Assignment01.
```{r load-data}
Assignment03DataRaw <- read.csv(here("Data/DataRaw/NBERPatCit95to99Sample.csv"))
```

## Transform Dependent Variable
The following code chunk creates a new variable called `CRECEIVEtr` which is a transformation of the `CRECEIVE` variable. I applied this transformation to create a new dependent variable because the histogram of the data for the `CRECEIVE` variable was positively skewed (i.e., skewed right).
```{r transform-DV}
Assignment03DataClean <- mutate(Assignment03DataRaw, CRECEIVErecip = (1/CRECEIVE))
```

## Save Data
The following code chunk saves the cleaned data.
```{r save-data}
write.csv(Assignment03DataClean, here("Data/DataClean/Assignment03DataClean.csv"), append = FALSE)
```
