---
title: "SOC 6100 Fall 2018 Assignment 01 R Notebook"
author: "Malcolm S. Townes"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
    github_document: default
    html_notebook: default
    always_allow_html: yes
---

## Introduction
This is an R Notebook for Assignment 01 in SOC 6100 Fall 2018.

## Project Set Up
The following code chunk enables the R Notebook to integrate seemlessly with the project organization format. This is normally included in the R Notebook to simplify file calls and enable file portability but it has been causing an error. Per Chris Prener of Saint Louis University, the error is generated because the `here::here()` function has not been tested with certain combinations of functions. To work around this problem, I've embedded the `here()` function where I enter a file path when necessary.

```{r setup-project}
knitr::opts_knit$set(root.dir = here::here())
```

## Load Dependencies
The following code chunk loads package dependencies required to perform the necessary tasks. Basic tasks include importing, reading, wrangling, and cleaning data; selecting a subset of the data; checking for unique observations, and analyzing missing data. 
```{r load-dependencies}
library(tidyverse) # loads the basic R packages
library(here) # enables file portability
library(readr) # functions for reading data
library(dplyr) # functions for data wrangling
library(janitor) # functions for data cleaning
library(naniar) # functions for analyzing missing data
```

## Load Raw Data
The following code chunk imports the raw data from the `CSV` file.
```{r import-data}
setwd("E:/SOC6100")
DataRaw <- read.table("Data/DataNBERPatentCitations/apat63_99/apat63_99.txt", sep = ",", header = TRUE, fill = TRUE, dec = ".") # import raw data
setwd("E:/SOC6100/Assignments/Assignment01")
```

## Subset Data
The following code chunk creates a subset for the data of interest.
```{r subset-data}
DataSubset <- filter(DataRaw, GYEAR>=1999) # subset data
DataSubset <- as_tibble(DataSubset) # convert data frame to tibble
```

## Inspect Data
The following code chunk evaluates the subsetted data.  It first checks for missing data for each variable.  It then checks for missing data for each variable in each observation.  Then it checks for duplicate observations with the `PATENT` variable to determine if that variable can be used as a unique identifier for each observation.  Finally, it checks for duplicate observations across all variables to ensure that each observation is unique.
```{r inspect-data}
miss_var_summary(DataSubset, order = TRUE)
miss_case_summary(DataSubset, order = TRUE)
get_dupes(DataSubset, PATENT)
get_dupes(DataSubset)
```

## Clean Data
I did not perform any additional cleaning of the subsetted data.
```{r view-data}
View(DataSubset, "Cleaned Data")
```

## Save Data
The following code chunk saves the cleaned data to a new file.
```{r save-data}
write.csv(DataSubset,here("Data/DataClean/NBERPatCit1999.csv"), append = FALSE)
```

